Effects of below-ground space limitation on performance of Eucalyptus seedlings:  Does photosynthesis really control growth?

### 2015-07-21 manuscript/manuscript_pve.Rmd
added tables to manuscript (jpg)
messed with srl stats and output
added srl changes in manuscript

Interpreting limitations to plant growth requires understanding of the balance between carbon (C) source and sink activity in order to assess C allocation and biomass partitioning. This study used manipulations of soil volume to test how growth is coupled to physiology, allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally sown seedlings (free). Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that soil volume effect would be largest in the smallest containers, resulting in physical constraints to growth independently of photosynthesis (A). Photosynthesis would then become sink-limited, resulting in the build-up of leaf nonstructural carbohydrates eventually leading to photosynthetic down regulation. We observed a negative container effect on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes mass, partitioning to leaves, stems and roots was conserved after 120 days. Photosynthetic capacity was also significantly reduced in containers, and was related to both leaf nitrogen content and starch accumulation. We developed a seedling growth model that utilized leaf A rates to allocate daily C uptake towards mass growth of stems, leaves and roots. We then asked whether the observed reductions in A explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in the down regulation of A, these reductions were not significant enough to explain observed growth responses. Thus, as photosynthesis and growth were not coordinated an excess pool of non-biomass C resulted in seedlings with soil volume restriction. This research highlights the need to further utilize mass balance approaches when evaluating plant C allocation and confirms that A and growth are not always directly related.

@@

\
SRL of harvested fine roots was not different across soil volumes (Table. 1). Over the duration of the experiment SLA was higher in free seedlings but was not different across containers sizes (Table. 1, P<0.001) and this pattern was evident in the first gas exchange measurement campaign (P<0.001).

@@

@@

## Conclusions
It has long been know that conditions which affect the photosynthetic process affect growth but that there is also an interrelationship between growth and A [@sweet1966role]. Here, the fate of assimilated C available for plant growth varied between naturally sown seedlings and seedlings with belowground sink inhibition. First, this is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles but how these results actually reflect field-grown plants has seldom been studied. Second, our finding add more evidence that A and growth are not always entirely synced, an important distinction often missed in studies that manipulate source/sink activity. Korner [@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control A through the demand for C  Although this may be true, we argue that attempts to quantify or at least predict the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents.  A lack of knowledge regarding C allocation restricts our ability to achieve mass balance and is a major obstacle in understanding the coordination between A and growth. In any case, we should not assume that

@@

>>>>

Interpreting limitations to plant growth requires understanding of the balance between carbon (C) source and sink activity in order to assess C allocation and biomass partitioning. This study used manipulations of soil volume to test how growth is coupled to physiology, allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally sown seedlings (free). Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that soil volume effect would be largest in the smallest containers, mediated by a down-regulation of photosynthesis (A). The accumulation of leaf non-structural carbohydrates, resulting from reduced sink strengh, was expected to correleltae to reductions in photosynthetic capacity. We observed a negative container effect on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes mass, partitioning to leaves, stems and roots was conserved after 120 days. Photosynthetic capacity was also significantly reduced in containers, and was related to both leaf nitrogen content and starch accumulation. We developed a seedling growth model that utilized leaf A rates to allocate daily C uptake towards mass growth of stems, leaves and roots. We then asked whether the observed reductions in A explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in the down regulation of A, these reductions were not significant enough to explain observed growth responses. Thus, as photosynthesis and growth were not coordinated an excess pool of non-biomass C resulted in seedlings with soil volume restriction. This research highlights the need to further utilize mass balance approaches when evaluating plant C allocation and confirms that A and growth are not always directly related.

@@

\
Overall, SRL was higher in seedlings in containers than free seedlings but not significantly in every soil volume treatment   (Table. 1). Over the duration of the experiment SLA was higher in free seedlings but was not different across containers sizes (Table. 1, P<0.001) and this pattern was evident in the first gas exchange measurement campaign (P<0.001).

@@

\
Alternatively, physical restriction of root proliferation could have impacted root development or morphology prior to shifts in mass partitioning. Here, increases in root length occurred in several of the soil volume treatments while no shifts in partitioning between fine and coarse root mass fractions were detected.  The overall poor quality of the soil or the effects of root restriction, however, likely decreased the capacity of this morphological response to increase leaf N uptake. This is not surprising, as plants in containers have been shown to have different root morphology to field grown plants as roots essentially compete with themselves for nutrients [@nesmith1998effect]. As the root mass fractions did not differ this provides evidence against an optimal foraging response in seedlings in containers. This could be because lateral root development is affected by inanimate obstacles and avoiding growth towards container walls could improve the efficiency of resource allocation [@falik2005root]. Also, the sensitivity of roots to their own exudates near obstructions may be used to adjust growth before stressful conditions occur [@semchenko2008foraging].

@@

## Conclusions
It has long been know that conditions which affect the photosynthetic process affect growth but that there is also an interrelationship between growth and A [@sweet1966role]. Here, the fate of assimilated C available for plant growth varied between naturally sown seedlings and seedlings with belowground sink inhibition. First, this is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles but how these results actually reflect field-grown plants has seldom been studied. Second, our finding add more evidence that A and growth are not always entirely synced, an important distinction often missed in studies that manipulate source/sink activity. Korner [@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control A through the demand for C  Although this may be true, we argue that attempts to quantify or at least predict the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents.  A lack of knowledge regarding C allocation restricts our ability to achieve mass balance and is a major obstacle in understanding the coordination between A and growth. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

@@

<!---
![table2](harvesttable.jpg)
![table2](phystable.jpg)
![table3](modeltable.jpg)
-->
